Comparison of dimensionality reduction schemes for derivative-free global optimization algorithms

### 2018-05-14 2018/YSC/ysc_paper.tex
YSC: fix running header
YSC: fix typos

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Comparison of dimensionality reduction schemes for dervative-free global optimization
algorithms}

@@

\titlerunning{Dimension reduction schemes in direct GO}

@@

In recent years several methods of constructing Peano-type space-filling curves (\textit{evolvents})
have been proposed \cite{strongin1978,Strongin1992,Goryachih2017,Gergel2009}. Some of them were
succesful used to build large-scale parallel algorithms \cite{stronginGergelBarkalovParGO}.
Presented paper numerically compares different evolvents in terms of convergence rates and
computational efficiency. The comparison is done by solving sets consisting of hundreds of test problems and collecting
different characteristics of convergence. As a result, practical recommendations about
the usage of each evolvent type are given. For example, some evolvents are suitable for parallel computations,
while others are quite useful in case of sequential solving of problems with time-consuming objective functions.

@@

The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the
described algorithm are given, for example, in \cite{strSergGO}.

@@

%------------------------------------------------------------------------------
\section{Conclusions}
In the present work, 5 different Peano curve-type mappings applied to the dimensionality
reduction in the global optimization problems were considered.
From the preliminary comparison conducted in Sec.~\ref{sec:seq_comp}, one can make the
following conclusions:
\begin{itemize}
  \item the smooth evolvent and the non-univalent one demonstrate the best result in the
problems of small dimensionality and can be applied successfully in solving the problems with
the computational costly objective functions. The properties of these evolvents donï¿½t allow
developing the optimization algorithms scalable onto several cluster nodes based on these ones.
  \item the shifted evolvents introduce large overhead costs on the operation of the method due
to the requirement to adding an auxiliary functional constraint into the problem (\ref{eq:task}).
The experiments have demonstrated that up to 95\% of the search information account for the
points, in which the auxiliary constraint is computed only.
However, if the objective function is computation-costly enough, the use of these evolvents
could make sense.
  \item the rotated evolvents have provided an acceptable speed of convergence in the problems
of small dimensionality in the sequential mode. The use of these ones doesn't result in the
introduction of the auxiliary constraints that allows constructing an efficient parallel algorithm
based on these evolvents.
\end{itemize}

>>>>

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Comparison of dimensionality reduction schemes for derivative-free global optimization
algorithms}

@@

\titlerunning{Dimension reduction schemes in derivative-free GO}

@@

In recent years several methods of constructing Peano-type space-filling curves (\textit{evolvents})
have been proposed \cite{strongin1978,Strongin1992,Goryachih2017,Gergel2009}. Some of them were
successful used to build large-scale parallel algorithms \cite{stronginGergelBarkalovParGO}.
Presented paper numerically compares different evolvents in terms of convergence rates and
computational efficiency. The comparison is done by solving sets consisting of hundreds of test problems and collecting
different characteristics of convergence. As a result, practical recommendations about
the usage of each evolvent type are given. For example, some evolvents are suitable for parallel computations,
while others are quite useful in case of sequential solving of problems with time-consuming objective functions.

@@

The convergence conditions and exact formulas for decision rules $R(i)$ and $d(x)$ of the
described algorithm are given, for example, in \cite{strSergGO}.

@@

%------------------------------------------------------------------------------
\section{Conclusions}
In the present work, 5 different Peano curve-type mappings applied to the dimensionality
reduction in the global optimization problems were considered.
From the preliminary comparison conducted in Sec.~\ref{sec:seq_comp}, one can make the
following conclusions:
\begin{itemize}
  \item the smooth evolvent and the non-univalent one demonstrate the best result in the
problems of small dimensionality and can be applied successfully in solving the problems with
the computational costly objective functions. The properties of these evolvents doesn't allow
developing the optimization algorithms scalable onto several cluster nodes based on these ones.
  \item the shifted evolvents introduce large overhead costs on the operation of the method due
to the requirement to adding an auxiliary functional constraint into the problem (\ref{eq:task}).
The experiments have demonstrated that up to 95\% of the search information account for the
points, in which the auxiliary constraint is computed only.
However, if the objective function is computation-costly enough, the use of these evolvents
could make sense.
  \item the rotated evolvents have provided an acceptable speed of convergence in the problems
of small dimensionality in the sequential mode. The use of these ones doesn't result in the
introduction of the auxiliary constraints that allows constructing an efficient parallel algorithm
based on these evolvents.
\end{itemize}
